---
title: "Metric Analysis Code"
format: html
execute: 
  echo: true
  warning: false
---

File created on 2025-10-06

# 0. Setup
Add any packages that are needed for analysis in this code chunk. 
```{r packages}
library(IEAnalyzeR)
```

File Naming Setup.
!! Auto generated-Do Not Change !!
```{r}

root_name<- "test_met"

csv_filename<-paste0("data/formatted/formatted_csvs/", root_name, "_formatted.csv") 
object_filename<-paste0("data/formatted/final_objects/", root_name, "_object.rds")
plot_filename<-paste0("figures/plots/", root_name, "_plot.png")

```

# 1. Read Data
Pull data from its source:
Manual data: data/unformatted data
Automated data: Add script for data call (API, package, etc.)
Confidential data: Store locally in the confidential data folder
    - This folder is excluded using gitignore and will not push to the GitHub repo
```{r}

```

# 2. Data Transformation
Here is where you transform the data to fit the IEA data format. 
For more info on IEA data format go HERE, or use THIS FUNCTION to view dummy data.
```{r}

```

# 3. Create Data_Prep object
Please use your formatted CSV to create a "data_prep" object.
For more info on the data_prep function please go HERE
```{r}
data_obj<-IEAnalyzeR::data_prep()
```

# 4. Save Formatted csv and data_prep object
This will save your data to the appropriate folders. 
Please replace "formatted_csv" with the name of your final formatted data.
```{r}
#Save formatted CSV
write.csv(formatted_csv, file = csv_filename, row.names = F)

#Save data_prep object
saveRDS(data_obj, file = object_filename)
```

# 5. Preview Plot
Use the IEAnalyzeR plotting function to preview the data. This will not necessarily be the final figure used in reports.
For more info on the plot_fn_obj function go HERE
```{r}

IEAnalyzeR::plot_fn_obj(df_obj = data_obj)

```

# 6. Save plot
This will save the plot to the correct folder.
Adjust height & width using (height=, width=, unit="in") if needed.
```{r}
ggsave(filename = plot_filename)
```

